---
title: "R Notebook for Monty Hall simulations"
author: "@jonaslau on Github"
output: html_notebook
---

# Introduction
This exercise creates a simulated answer to the [Monty Hall](https://en.wikipedia.org/wiki/Monty_Hall_problem). 

## Import tidyverse package
```{r, warning=FALSE}
library(tidyverse)
```

## Create some global variables, set seed
```{r}
set.seed(8888)

# number of simulated games
n_games = 5000
```

## Monty Hall data frame
```{r}
monty_hall = tibble(
  prize_room = sample(c("a","b","c"), size = n_games, replace = TRUE), # room with the prize, random sampled
  first_guess = sample(c("a","b","c"), size = n_games, replace = TRUE), # first guess by player, random sampled
  open = NA, # door opened by host after first guess
  switch = sample(c(0,1), size=n_games, replace = TRUE), # whether the player switch door after a door opened, random sampled
  final_choice = NA, # final response by the player
  win_firstguess = NA, # winning for the first guess: placeholder
  win_end = NA # winning at the end: placeholder
)
```

## Determine which door the host can open
```{r}
# open door
monty_hall = monty_hall %>%
  mutate(open = case_when(
    # the host's choices for opening a door
    # e.g., if prize is behind door a, and the player chose door a
    # the host can only open door b or door c
    prize_room == "a" & first_guess == "a" ~ sample(c("b", "c"), size = 1),
    prize_room == "a" & first_guess == "b" ~ "c",
    prize_room == "a" & first_guess == "c" ~ "b",
    # same logic with prize behind door b
    prize_room == "b" & first_guess == "b" ~ sample(c("a", "c"), size = 1),
    prize_room == "b" & first_guess == "a" ~ "c",
    prize_room == "b" & first_guess == "c" ~ "a",
    # same logic with prize behind door c
    prize_room == "c" & first_guess == "c" ~ sample(c("a", "b"), size = 1),
    prize_room == "c" & first_guess == "a" ~ "b",
    prize_room == "c" & first_guess == "b" ~ "a",
    TRUE ~ NA_character_)
)
```

## Determine the final guess
Two conditions: the player changes / does not change mind
```{r}
monty_hall = monty_hall %>%
  mutate(final_choice = case_when( 
    switch == 0 ~ first_guess, # if player does not change mind, final choice is the first guess
    switch == 1 ~ case_when(
      # if the player change mind, there is only one choice left
      first_guess == "a" & open == "b" ~ "c",
      first_guess == "a" & open == "c" ~ "b",
      first_guess == "b" & open == "a" ~ "c",
      first_guess == "b" & open == "c" ~ "a",
      first_guess == "c" & open == "a" ~ "b",
      first_guess == "c" & open == "b" ~ "a",
    )
  ))
```

## Calculate the overall accuracy if players switch/ not switch
```{r}
# calculate whether win or not
monty_hall = monty_hall %>%
  mutate(
    win_firstguess = as.integer(prize_room == first_guess), # convert binary var to numbers
    win_end = as.integer(prize_room == final_choice)
  )
```

## Some winning statistics
```{r}
# statistics for first guess
firstguess = monty_hall %>%
  pull(win_firstguess) %>%
  mean() * 100

sprintf("A random guess would yield a %.1f%% accuracy", firstguess)

# average for winning, grouped by the decision to switch
winning = monty_hall %>%
  group_by(switch) %>%
  summarise(mean_accuracy = mean(win_end) * 100)

sprintf("Without switching, chance of winning is %.1f%%; with switching, chance of winning becomes %.1f%%", 
        winning %>% filter(switch == 0) %>% pull(mean_accuracy), 
        winning %>% filter(switch == 1) %>% pull(mean_accuracy))
```